---
title: "Investigate the Exponential Distribution in R"
author: "jbratanov"
date: "Monday, May 23, 2016"
output: html_document
---

####We will investigate the exponential distribution using R and compare it with the Central Limit Theorem.  
####The following criteria will be used for this project:
 1. mean of exponential distribution = 1/lambda
 2. standard deviation = 1/lambda
 3. lambda = 0.2 for simulations
 4. investigate distribution of 40 exponentials (will need to do 1000 simulations)

### Compare sample mean to the theoretical mean of the distribution
```{r}

# simulation vars
lambdaVal <- .2
simCnt <- 1000
n <- 40

# set seed to 1.  Easy to remember
set.seed(1)

# Run simulation to gather data
simMatrix <- matrix(rexp(simCnt * n, rate=lambdaVal), simCnt, n)
simData <- rowMeans(simMatrix)

# Get simulated mean
simMean <- mean(simData)

# Get theoretical mean
theoryMean <- 1 / lambdaVal

# Print mean comparison
cat("Sample mean =", simMean, "and Theoretical mean =", theoryMean)
```
####The sample mean of 4.99 is very close to theoretical mean of 5.

### Compare sample variance to the theoretical variance of the distribution
```{r}

# Get simulated variance
simVar <- var(simData)

# Get theoretical variance
theoryVar <- (1/lambdaVal)^2/n

# Print variance comparison
cat("Sample variance =", simVar, "and Theoretical variance =", theoryVar)
```
####The sample variance of 0.6177 is very close to theoretical variance of 0.625.

### Show that the distribution is approximately normal
```{r}



```




