---
title: "Investigate the Exponential Distribution in R"
author: "jbratanov"
date: "Monday, May 23, 2016"
output: html_document
---

####We will investigate the exponential distribution using R and compare it with the Central Limit Theorem.  
####The following criteria will be used for this project:
 1. mean of exponential distribution = 1/lambda
 2. standard deviation = 1/lambda
 3. lambda = 0.2 for simulations
 4. investigate distribution of 40 exponentials (will need to do 1000 simulations)

## Compare sample mean to the theoretical mean of the distribution
```{r}

# constant variable values for simulation
# lambda is the rate
# n is the sample size
# simCnt is the number of simulations
lambdaVal <- .2
simCnt <- 1000
n <- 40

# set seed to 1.  Easy to remember
set.seed(1)

# Run simulation to gather data
# use rexp for random generation of the exponential distribution
# store the simulation data in a matrix 
simMatrix <- matrix(rexp(simCnt * n, rate=lambdaVal), simCnt, n)
simData <- rowMeans(simMatrix)

# Calculate the simulated mean
simMean <- mean(simData)

# Calculate the theoretical mean
theoryMean <- 1 / lambdaVal

# Print mean comparison
cat("Sample mean =", simMean, "and Theoretical mean =", theoryMean)
```
####The sample mean of 4.99 is very close to theoretical mean of 5.

## Compare sample variance to the theoretical variance of the distribution
```{r}

# Calculate the simulated variance
simVar <- var(simData)

# Calculate the theoretical variance
theoryVar <- (1/lambdaVal)^2/n

# Print variance comparison
cat("Sample variance =", simVar, "and Theoretical variance =", theoryVar)
```
####The sample variance of 0.6177 is very close to theoretical variance of 0.625.

## Show that the distribution is approximately normal
```{r}
library(ggplot2)
# Need to create a data frame for ggplot to use
simData_df <- data.frame(simData)

# Create graph to show normal distribution
# ggplot will build histogram using simData
# We will use stat_function to display the normal distribution to compare
ggplot(simData_df, aes(x = simData)) + 
    geom_histogram(aes(y=..density..), fill = I('lightblue'), 
                   binwidth = 0.20, color = I('black')) + 
  xlab("Exponential Distribution of Simulation Data for Lambda .2") +
  ylab("Frequency") +
  stat_function(fun = dnorm, args = list(mean = 5, sd = sd(simData_df$simData)))
```

####  The above graph simulated data is approximately distributed to the normal mean.
